Poincar\'e--Friedrichs constants over domains with shellable triangulations

### 2024-07-04 main.tex
Update on Overleaf.

\begin{abstract}
    We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    The main application that we envision are computable estimates over finite element patches. 
    %
    As a result with independent standing, we establishes Poincar\'e--Friedrichs inequalities for the $L^{p}$ de~Rham complex over convex bounded domains. 
    This is achieved through bounding operator norms of Poincar\'e and Bogovski\u{\i} potential operators. 
\end{abstract}

@@

For example, if $\bff \in \bfL^{2}(\Omega)$ is a square-integrable vector field on some domain $\Omega \subseteq \bbR^3$, 
we want to find a potential $\bfu \in \bfL^2(\Omega)$ under the curl operator whose $\bfL^2$ norm is bounded in terms of the source field. This is expressed via the Poincar\'e--Friedrichs inequality 
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfL^{2}(\Omega) \\ \curl \bfv = \bff } } \| \bfv \|_{\bfL^{2}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{\bfL^{2}(\Omega)}
    .
\end{align*}
The constant in this inequality quantifies the stability properties of the Maxwell system on that domain. Any upper bound for it corresponds to a lower bound for the Maxwell eigenvalues.

Finding a potential operator, whether linear or non-linear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
% We accomplish our results for a special class of triangulations, so-called shellable triangulations.
We accomplish this undertaking in the special case where $\Omega$ admits a so-called \emph{shellable} triangulation. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards the more general case. 
For example, local patches around a simplex within a larger 2D or 3D triangulation are shellable.
We express all our main results in the calculus of differential forms, also known as exterior calculus, 
and treat the curl and divergence operator as its particular instances. 
\\

@@

Most research on optimal constants has addressed the case of convex domains. 
% The optimal constants that depend solely on the domain diameter are known if $p=2$ \cite{bebendorf2003note} and if $p=1$ \cite{acosta2004optimal}. 
Notably, the optimal constants $C_{\grad,p}$ for any $1 \leq p < \infty$ that depends only on the domain diameter is known when the domain is convex.~\cite{bebendorf2003note,acosta2004optimal,esposito2013poincare,ferone2012remark}.
When the domain is not convex but star-shaped with respect to a ball,
then polynomial interpolation estimates already imply the Poincar\'e inequality~\cite{ern2021finite}. 
An open set star-shaped with respect to a internal point is also star-shaped with respect to some open ball within the domain,
implying Poincar\'e inequalities on star-shaped open sets~\cite{hurri1988poincare}.

The situation is less clear cut when we study the curl or divergence operators in vector calculus.
We use the vector field spaces
\begin{align*}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \curl \bfu \in \bfL^p(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \divergence \bfu \in L^p(\Omega) \right\}
    .
\end{align*}
Their members are vector fields whose distributional partial derivatives satisfy certain cancellation properties. 
Our objective is to find bounded potentials for the operators 
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow \bfL^{p}(\curl,\Omega),
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega),
\end{align*}
The fundamental difference is that the curl and divergence have infinite-dimensional kernels. 
In the Banach space case, it is not immediately evident that the kernels of these operators are complemented.
The existence of a bounded linear potential operator is a non-trivial fact even in the Hilbert space case.
We are interested in constants for inequalities such as 
\begin{align*}
    \argmin\limits_{ \substack{ \bfv \in \bfW^{p}(\curl,\Omega) \\ \curl \bfv = \bff } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\curl,p}
    \| \bff \|_{\bfL^{p}(\Omega)}
    ,
    \qquad 
    \argmin\limits_{ \substack{ \bfv \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfv = f } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\divergence,p}
    \| f \|_{L^{p}(\Omega)}
    ,
\end{align*}
and whether these constants can be estimated by the operator norms of linear potential operators 
\begin{align*}
    \Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \quad 
    \Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
Not much attention seems to have been given to the study of optimal constants in such Poincar\'e--Friedrichs inequalities for the curl and divergence operators. 
% even if the domain is convex, 
This sharply contrasts to the study of the gradient operator.

The literature is much scarcer when it comes to this. 
We can access some estimates from the literature in the formalism of exterior calculus. 
We are aware of work by Guerini and Savo~\cite{guerini2004eigenvalue}, who estimate the Poincar\'e--Friedrichs constants over convex domains in the Hilbert space setting, when $p=2$. They provide lower bounds for the spectrum of the Hodge--Laplace operator on convex domains. This implies upper bounds for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence for the curl and divergence operators.
By duality, that implies lower bounds for the case of full boundary conditions.
Their work also includes the following sequence of inequalities,
valid over convex domains in 3D:
\begin{align}\label{math:intro:guerinisavo}
    C_{\grad,2} \geq C_{\curl,2} \geq C_{\divergence,2}.
\end{align}
When the domain is star-shaped with respect to a ball, then estimates for the Poincar\'e--Friedrichs constants 
follow from bounds on the operator norms of regularized Poincar\'e operators~\cite{costabel2010bogovskiui}
as mappings between Lebesgue spaces, though these estimates have not been made explicit in the literature yet.
\\

Clearly, not all local patches describe convex subdomains. 
One possible avenue towards Poincar\'e--Friedrichs constants is the observation that local patches in triangulations are often star-shaped with respect to a ball. 
Over patches around interior subsimplices, the size of the interior ball only depends on the shape regularity of the triangulation. 
However, that interior ball can be arbitrarily small when the patch is around a boundary simplex at a re-entrant corner, even if the mesh has good shape regularity. 
Estimates that rely on this geometric condition deteriorate when the aforementioned ball has small radius,
and that is inevitable when the domain has an unfortunate geometry. 
Some illustrative limit cases include the slit domain and the crossed bricks domain~\cite{licht2019smoothed},
for which some finite element patches are not star-shaped with respect to any ball.

Domains that admit finite triangulations allow for different techniques to attain Poincar\'e--Friedrichs inequalities. 
However, the majority of results only address such inequalities for scalar-valued functions, not for vector fields. 
We review some of the main outcomes. 
% TODO: wait for answer to email 
Veeser and Verf\"urth provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for finite element stars~\cite{veeser2012poincare}. Naturally, their estimate deteriorates when the shape regularity of the mesh deteriorates. 
% 
The literature knows different variations of a popular avenue towards Poincar\'e--Friedrichs constants that can circumvent the effect of low boundary regularity~\cite{vohralik2005discrete,ern2021finite}. 
These potentials share being constructed step-by-step: 
starting with a single triangle, the potential is constructed over increasing subdomains, until the entire domain is complete. That procedure applies to general triangulated domains, not only local stars, though the latter is our main interest. 
The underlying idea is that we first construct a potential for the gradient over an initial simplex. 
Every time we have found a potential over a subdomain, we construct a potential over a neighboring simplex or patch:
along the interfacing intersection, the two potentials will only differ by a constant, 
and that difference can easily be removed to ensure continuity across that interface. 
% Estimates for Poincar\'e--Friedrichs constants of triangulated domains are known in various forms.

While we know Poincar\'e--Friedrichs constants over finite element patches for scalar-valued Sobolev spaces, 
we are not aware of computable estimates for the case of $\bfW^{p}(\curl,\Omega)$ and $\bfW^{p}(\divergence,\Omega)$ over finite element stars. 
In the Hilbert space setting, with $p=2$, the Friedrichs inequality over the Sobolev space with Dirichlet boundary conditions implies, 
via duality, the Poincar\'e--Friedrichs for the divergence operator. However, that easy observation seems restricted to the Hilbert space case $p=2$. 
We are not aware of any estimates for the curl operator, even in the Hilbert space situation. 
\\

The main objective of the present exposition is the construction of linear potential operators the gradient, curl, and divergence, satisfying explicitly computable upper bounds. We follow closely the philosophy of the aforementioned sequential approach. Transitioning to vector fields incurs new challenges due to the infinite-dimensional kernels. We overcome these challenges on a special class of triangulated domains, namely domains with so-called \emph{shellable} triangulations.

@@

In addition to that, we need to find potentials over a simplex subject to partial boundary conditions,
Again, we want to address the entire $1 \leq p \leq \infty$ of Lebesgue exponents. 
We are not aware of any explicit estimates in the literature and there are no regularized potential operators for these boundary conditions. 
The Poincar\'e--Friedrichs inequality subject to full boundary conditions is repurposed to prove Poincar\'e--Friedrichs inequalities for differential forms over simplices and subject to partial boundary conditions. For example, given the curl of a vector field over a tetrahedron that vanishes along, say, two of the tetrahedron faces, we want to find the unique vector field potential. We achieve this by constructing an auxiliary problem subject to full boundary conditions.
\\

% Future work 
Computable reliable estimates for the Poincar\'e--Friedrichs constant are important: 
they quantify the fundamental stability properties of partial differential equations in vector analysis,
such as the Hodge Laplace equation. 
Upper bounds correspond to lower bounds for the non-zero eigenvalues of the scalar and vector Laplacian.
Furthermore, their constants enter the stability and convergence theory of finite element methods. 
In numerical methods, upper bounds for the Poincar\'e--Friedrichs constant enter the stability and convergence theory of finite element methods,
and they lead to lower bounds for the spectrum of finite element matrices.

@@

We have also identified that the Guerini--Savo estimates, hitherto only known for the exterior derivative as a mapping between Hilbert spaces, could be extended to the entire range of Lebesgue exponents $1 \leq p \leq \infty$. We believe such estimates would yield considerably better bounds on the Poincar\'e--Friedrichs constants than the regularized potential operators. It is of interest whether the Poincar\'e--Friedrichs constant for the gradient operator still bounds the Poincar\'e--Friedrichs constants of the other instances of the exterior derivatives, as in \eqref{math:intro:guerinisavo}, when $1 \leq p \leq \infty$ and the domain is convex. On a related note, explicit estimates for the operator norms of regularized potential operators with different degrees of smoothing are of general interest. These topics remain for future research. 
\\

@@

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation}, if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
% (i) every subsimplex of any simplex in $\calT$ is itself part of $\calT$, and (ii) the intersection between any two simplices $T, T' \in \calT$ is either empty or a shared subsimplex of both $T$ and $T'$. 
The set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that complex.\footnote{We warn the reader that simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes.} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, which is defined as the union 
\begin{align*}
    \underlying{\calT} := \bigcup_{ T \in \calT } T
    .
\end{align*}
We call a set \textit{trianguable} if it is the underlying set of some triangulation.

@@

\begin{lemma}
 Let $\calT$ be an $n$-dimensional simplicial complex and let $S,S' \in \calT$.
 Then either $\calT_{n}(S) \cap \calT_{n}(S') = \emptyset$ or there exists $S'' \in \calT$
 such that $\calT_{n}(S) \cap \calT_{n}(S') = \calT_{n}(S'')$ with $S, S' \subseteq S''$.
\end{lemma}
\begin{proof}
 Let $T \in \calT$ be $n$-dimensional.
 We have $T \in \patch_{\calT}(S )$ if and only if all vertices of $S $ are vertices of $T$.
 We have $T \in \patch_{\calT}(S')$ if and only if all vertices of $S'$ are vertices of $T$.
 Consequently, $T \in \patch_{\calT}(S) \cap \patch_{\calT}(S')$ if and only if $T \in \patch_{\calT}(S'')$,
 where $S'' \in \calT$ is the convex closure of $S$ and $S'$.
\end{proof}

@@

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature. Given $p \in [1,\infty]$ and an open set $\Omega \subseteq \bbR^{n}$ of finite measure, 
we say that $\Omega$ satisfies the Poincar\'e inequality with exponent $p$ 
if there exists $C_{{\varnothing},\Omega,p} \geq 0$ such that 
for all $u \in W^{1,p}(\Omega)$
\begin{align}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} 
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
\end{align}
where $u_{\Omega}$ is the \emph{average} of $u$ over $\Omega$, that is,
\begin{align*}
    u_{\Omega} := |\Omega|^{-1} \int_{\Omega} u.
\end{align*}
Clearly, this Poincar\'e inequality implies the Poincar\'e--Friedrichs inequality and we have $C_{{\rm{PF}},\Omega,p} \leq C_{\varnothing,\Omega,p}$. 
Towards a converse inequality, 
let us first observe that the average of any $v \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align*}
    \| v_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)| \right)^{p}
    \leq 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)|^{p} \right)
    = 
    \| v \|_{L^{p}(\Omega)}^{p}
    .
\end{align*}
Here, we have used Jensen's inequality. 
In the case $p = \infty$, any $v \in L^{\infty}(\Omega)$ satisfies $\| v_\Omega \|_{L^{\infty}(\Omega)} \leq \| v \|_{L^{\infty}(\Omega)}$. 
It follows that 
\begin{align*}
    \| v - v_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| v \|_{L^{p}(\Omega)},
    \quad 
    v \in L^{p}(\Omega)
    .
\end{align*}
% \begin{align*}
%     \color{Emerald}
%     \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p-1} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     +
%     2^{p-1} 
%     \| u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     .
% \end{align*}
Hence the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with $C_{\varnothing,\Omega,p} \leq 2 C_{{\rm PF},\Omega,p}$.

>>>>

\begin{abstract}
    We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    The main application that we envision are computable estimates over finite element patches. 
\end{abstract}

@@

For example, if $\bff \in \bfL^{2}(\Omega)$ is a square-integrable vector field on some domain $\Omega \subseteq \bbR^3$, 
we want to find a potential $\bfu \in \bfL^2(\Omega)$ under the curl operator whose $\bfL^2$ norm is bounded in terms of the source field. This is expressed via the Poincar\'e--Friedrichs inequality 
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfL^{2}(\Omega) \\ \curl \bfv = \bff } } \| \bfv \|_{\bfL^{2}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{\bfL^{2}(\Omega)}
    .
\end{align*}
The constant in this inequality quantifies the stability properties of the Maxwell system on that domain. Any upper bound for it gives rise to a lower bound for the Maxwell eigenvalues.

Finding a potential operator, whether linear or non-linear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
% We accomplish our results for a special class of triangulations, so-called shellable triangulations.
We accomplish this task in the special case where $\Omega$ admits a so-called \emph{shellable} triangulation. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards the more general case. 
For example, local patches around a simplex within a larger 2D or 3D triangulation are shellable.
We express all our main results in the calculus of differential forms, also known as exterior calculus, 
and treat the curl and divergence operator as its particular instances. 
\\

@@

Most research on optimal constants has addressed the case of convex domains. 
% The optimal constants that depend solely on the domain diameter are known if $p=2$ \cite{bebendorf2003note} and if $p=1$ \cite{acosta2004optimal}. 
Notably, the optimal constants $C_{\grad,p}$ for any $1 \leq p < \infty$ that depends only on the domain diameter is known when the domain is convex.~\cite{bebendorf2003note,acosta2004optimal,esposito2013poincare,ferone2012remark}.
When the domain is not convex but star-shaped with respect to a ball,
then polynomial interpolation estimates already imply the Poincar\'e inequality~\cite{brenner2008finite,ern2021finite}. 
An open set star-shaped with respect to a internal point is also star-shaped with respect to some open ball within the domain,
implying Poincar\'e inequalities on star-shaped open sets~\cite{hurri1988poincare}.

The situation is less clear cut when we study the curl or divergence operators in vector calculus.
We use the vector field spaces
\begin{align*}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \curl \bfu \in \bfL^p(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \divergence \bfu \in L^p(\Omega) \right\}
    .
\end{align*}
Their members are vector fields whose distributional curls and divergences are in Lebesgue spaces. 
This only requires that certain linear combinations of distributional partial derivatives are measurable, 
and in particular these spaces are not classical Sobolev spaces of vector fields. 
Our objective is to find bounded potentials for the operators 
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow \bfL^{p}(\Omega),
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega),
\end{align*}
The fundamental difference is that the curl and divergence have infinite-dimensional kernels. 
In the Banach space case, it is not immediately evident that the kernels of these operators are complemented.
The existence of a bounded linear potential operator is a non-trivial fact even in the Hilbert space case.
We are interested in constants for inequalities such as 
\begin{align*}
    \argmin\limits_{ \substack{ \bfv \in \bfW^{p}(\curl,\Omega) \\ \curl \bfv = \bff } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\curl,p}
    \| \bff \|_{\bfL^{p}(\Omega)}
    ,
    \qquad 
    \argmin\limits_{ \substack{ \bfv \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfv = f } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\divergence,p}
    \| f \|_{L^{p}(\Omega)}
    ,
\end{align*}
and whether these constants can be estimated by the operator norms of linear potential operators 
\begin{align*}
    \Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \quad 
    \Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
Not much attention seems to have been given to the study of optimal constants in such Poincar\'e--Friedrichs inequalities for the curl and divergence operators. 
% even if the domain is convex, 
This sharply contrasts to the study of the gradient operator.

The literature in functional analysis is much scarcer when it comes to this. 
We can access some estimates from the literature in the formalism of exterior calculus. 
We are aware of work by Guerini and Savo~\cite{guerini2004eigenvalue}, who estimate the Poincar\'e--Friedrichs constants over convex domains in the Hilbert space setting, when $p=2$. They provide lower bounds for the spectrum of the Hodge--Laplace operator on convex domains. This implies upper bounds for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence for the curl and divergence operators.
By duality, that implies lower bounds for the case of full boundary conditions.
Their work also includes the following sequence of inequalities,
valid over convex domains in 3D:
\begin{align}\label{math:intro:guerinisavo}
    C_{\grad,2} \geq C_{\curl,2} \geq C_{\divergence,2}.
\end{align}
When the domain is star-shaped with respect to a ball, then estimates for the Poincar\'e--Friedrichs constants 
follow from bounds on the operator norms of regularized Poincar\'e operators~\cite{costabel2010bogovskiui}
as mappings between Lebesgue spaces, though these estimates have not been made explicit in the literature yet.
\\

Clearly, not all local patches describe convex subdomains. 
One possible avenue towards Poincar\'e--Friedrichs constants is the observation that local patches in triangulations are often star-shaped with respect to a ball. 
Over patches around interior subsimplices, the size of the interior ball only depends on the shape regularity of the triangulation. 
However, that interior ball can be arbitrarily small when the patch is around a boundary simplex at a re-entrant corner, even if the mesh has good shape regularity. 
Estimates that rely on this geometric condition thus deteriorate when the aforementioned ball has small radius,
and that is inevitable when the domain has an unfortunate geometry. 
Some illustrative limit cases include the slit domain and the crossed bricks domain~\cite{licht2019smoothed},
for which some finite element patches are not star-shaped with respect to any ball.

Domains that admit finite triangulations allow for different techniques to attain Poincar\'e--Friedrichs inequalities. 
However, the majority of results only address such inequalities for scalar-valued functions, not for vector fields. 
We review some of the main outcomes. 
% TODO: wait for answer to email 
Veeser and Verfürth provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for finite element stars~\cite{veeser2012poincare}. Naturally, their estimate deteriorates when the shape regularity of the mesh deteriorates. 
% 
The literature knows different variations of a popular avenue towards Poincar\'e--Friedrichs constants that can circumvent the effect of low boundary regularity~\cite{vohralik2005discrete,ern2021finite}. 
These potentials share being constructed step-by-step: 
starting with a single triangle, the potential is constructed over increasing subdomains, until the entire domain is complete. That procedure applies to general triangulated domains, not only local stars, though the latter is our main interest. 
The underlying idea is that we first construct a potential for the gradient over an initial simplex. 
Every time we have found a potential over a subdomain, we construct a potential over a neighboring simplex or patch:
along the interfacing intersection, the two potentials will only differ by a constant, 
and that difference can easily be removed to ensure continuity across that interface. 
% Estimates for Poincar\'e--Friedrichs constants of triangulated domains are known in various forms.

While we know Poincar\'e--Friedrichs constants over finite element patches for scalar-valued Sobolev spaces, 
we are not aware of computable estimates for the case of $\bfW^{p}(\curl,\Omega)$ and $\bfW^{p}(\divergence,\Omega)$ over finite element stars. 
In the Hilbert space setting, with $p=2$, the Friedrichs inequality~\cite{burenkov1998sobolev} over the Sobolev space with Dirichlet boundary conditions implies, 
via duality, the Poincar\'e--Friedrichs inequality for the divergence operator. However, that easy observation seems restricted to the Hilbert space case $p=2$. 
We also remark that recent work~\cite{gallistl2023computational} addresses eigenvalues for the curl--curl operator over general triangulations,
which gives rise to an estimate for the Poincar\'e--Friedrichs constant when $p=2$. However, the use case of these estimates are sequences of triangulations after sufficiently many levels of refinement, whereas our main application are estimates that work with fixed but small triangulations, such as finite element stars. 
We are not aware of any estimates for the curl operator, even in the Hilbert space situation. 
\\

The main objective of the present exposition is the construction of linear potential operators for the gradient, curl, and divergence, satisfying explicitly computable upper bounds. We follow closely the philosophy of the aforementioned sequential approach. Transitioning to vector fields incurs new challenges due to the infinite-dimensional kernels. We overcome these challenges on a special class of triangulated domains, namely domains with so-called \emph{shellable} triangulations.

@@

In addition to that, we need to find potentials over a simplex subject to partial boundary conditions. 
Again, we want to address the entire $1 \leq p \leq \infty$ of Lebesgue exponents. 
We are not aware of any explicit estimates in the literature and there are no regularized potential operators for these boundary conditions. 
The Poincar\'e--Friedrichs inequality subject to full boundary conditions is repurposed to prove Poincar\'e--Friedrichs inequalities for differential forms over simplices and subject to partial boundary conditions. For example, given the curl of a vector field over a tetrahedron that vanishes along, say, two of the tetrahedron faces, we want to find the unique vector field potential. We achieve this by constructing an auxiliary problem subject to full boundary conditions.
\\

% Future work 
Computable reliable estimates for the Poincar\'e--Friedrichs constant are important: 
they quantify the fundamental stability properties of partial differential equations in vector analysis,
such as the Hodge Laplace equation. 
Upper bounds correspond to lower bounds for the non-zero eigenvalues of the scalar and vector Laplacian.
% Furthermore, their constants enter the stability and convergence theory of finite element methods. 
In numerical methods, upper bounds for the Poincar\'e--Friedrichs constant enter the stability and convergence theory of finite element methods,
and they lead to lower bounds for the spectrum of finite element matrices.

@@

We have also identified that the Guerini--Savo estimates, hitherto only known for the exterior derivative as a mapping between Hilbert spaces, could be extended to the entire range of Lebesgue exponents $1 \leq p \leq \infty$. We believe such estimates would yield considerably better bounds on the Poincar\'e--Friedrichs constants than the regularlized potential operators. It is of interest whether the Poincar\'e--Friedrichs constant for the gradient operator still bounds the Poincar\'e--Friedrichs constants of the other instances of the exterior derivatives, as in \eqref{math:intro:guerinisavo}, when $1 \leq p \leq \infty$ and the domain is convex. On a related note, explicit estimates for the operator norms of regularized potential operators with different degrees of smoothing are of general interest. These topics remain for future research. 
\\

@@

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation}, if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
% (i) every subsimplex of any simplex in $\calT$ is itself part of $\calT$, and (ii) the intersection between any two simplices $T, T' \in \calT$ is either empty or a shared subsimplex of both $T$ and $T'$. 
% The set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that complex.\footnote{We warn the reader that simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes.} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, which is defined as the union 
\begin{align*}
    \underlying{\calT} := \bigcup_{ T \in \calT } T
    .
\end{align*}
We call a set \textit{trianguable} if it is the underlying set of some triangulation.

@@

\begin{lemma}
 Let $\calT$ be an $n$-dimensional simplicial complex and let $S,S' \in \calT$.
 Then either $\patch_{\calT}(S) \cap \patch_{\calT}(S') = \emptyset$ or there exists $S'' \in \calT$
 such that $\patch_{\calT}(S) \cap \patch_{\calT}(S') = \calT_{n}(S'')$ with $S, S' \subseteq S''$.
\end{lemma}
\begin{proof}
 Let $T \in \calT$ be $n$-dimensional.
 We have $T \in \patch_{\calT}(S )$ if and only if all vertices of $S $ are vertices of $T$.
 We have $T \in \patch_{\calT}(S')$ if and only if all vertices of $S'$ are vertices of $T$.
 Consequently, $T \in \patch_{\calT}(S) \cap \patch_{\calT}(S')$ if and only if $T \in \patch_{\calT}(S'')$,
 where $S'' \in \calT$ is the convex closure of $S$ and $S'$.
\end{proof}

@@

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature. Given $p \in [1,\infty]$ and an open set $\Omega \subseteq \bbR^{n}$ of finite measure, 
we say that $\Omega$ satisfies the Poincar\'e inequality with exponent $p$ 
if there exists $C_{{\varnothing},\Omega,p} \geq 0$ such that 
for all $u \in W^{1,p}(\Omega)$
\begin{align}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} 
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
\end{align}
where $u_{\Omega}$ is the \emph{average} of $u$ over $\Omega$, that is,
\begin{align*}
    u_{\Omega} := |\Omega|^{-1} \int_{\Omega} u.
\end{align*}
Clearly, this Poincar\'e inequality implies the Poincar\'e--Friedrichs inequality and we have $C_{{\rm{PF}},\Omega,p} \leq C_{\varnothing,\Omega,p}$. 
Towards a converse inequality, 
let us first observe that the average of any $v \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align*}
    \| v_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)| \right)^{p}
    \leq 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)|^{p} \right)
    = 
    \| v \|_{L^{p}(\Omega)}^{p}
    .
\end{align*}
Here, we have used Jensen's inequality. 
In the case $p = \infty$, any $v \in L^{\infty}(\Omega)$ satisfies $\| v_\Omega \|_{L^{\infty}(\Omega)} \leq \| v \|_{L^{\infty}(\Omega)}$. 
It follows that 
\begin{align*}
    \| v - v_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| v \|_{L^{p}(\Omega)},
    \quad 
    v \in L^{p}(\Omega)
    .
\end{align*}
% \begin{align*}
%     \color{Emerald}
%     \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p-1} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     +
%     2^{p-1} 
%     \| u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     .
% \end{align*}
Via an application of the triangle inequality, 
the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with $C_{\varnothing,\Omega,p} \leq 2 C_{{\rm PF},\Omega,p}$.

In the special case $p=2$, we always have $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $v \in L^{2}(\Omega)$,
and therefore even $C_{\varnothing,\Omega,2} = C_{{\rm PF},\Omega,2}$. 
This follows from the orthogonality constant functions to average-free functions in $L^2(\Omega)$,
and it also follows from the projection estimate (see, e.g.,~\cite{xu2003some}).
%\footnote{\color{red}Is there a better name for this than Xu-Zikatanov estimate?}
Stern~\cite{stern2015banach} provided a generalization of the projection estimate that implies improved estimate for all $L^p$ spaces with $1 \leq p \leq \infty$:
since taking the average is a the projection onto the constants functions with unit norm, it now follows that % TODO: cite specific results of Ari 
\begin{align*}
    \| v - v_\Omega \|_{L^{p}(\Omega)}
    \leq 
    \min\left( 2, 2^{|2/p-1|} \right)
    \| v \|_{L^{p}(\Omega)}
    = 
    2^{|2/p-1|} 
    \| v \|_{L^{p}(\Omega)}
    ,
    \quad 
    v \in L^p(\Omega)
    .
\end{align*}
We have $1 \leq 2^{|2/p-1|} \leq 2$ for $1 \leq p \leq \infty$.
The upper bound is only attained in the limit cases $p = 1$ and $p = \infty$, and the lower bound is only attained if $p = 2$, which reproduces the improved estimate for the Hilbert space case.
The main consequence for us is the estimate 
$C_{\varnothing,\Omega,p} \leq 2^{|2/p-1|} C_{{\rm PF},\Omega,p}$.
We conclude that our notion of Poincar\'e--Friedrichs constant is equivalent to the notion of Poincar\'e constant in the literature, up to a numerical factor that only depends on $1 \leq p \leq \infty$.
